### 2018-04-11 significance/significance.Rmd

---
title: "Correspondence in Significance"
output: pdf_document
---

Programs of research that examine the replicability of research findings have become more common in psychology. 
However, there is some ambiguity over how to assess the findings of such programs; that is, how do we know if a finding has been replicated?
The Open Science Collaboration (2015) argue, "There is no single standard for evaluating replication success," and then analyzed replicate pairs in a variety of ways.
Subsequent programs have used some of these same analysis methods (e.g., Camerer, et al., 2016; Schweinsberg, et al., 2016).

It is not entirely clear what these methods say about replication.
For one, an actual definition of replication remains somewhat elusive---or flexible---across these methods. 
Saying that studies get "the same result" is somewhat subjective, and the OSC attempted to operationalize this by describing several different scenarios of studies getting "the same result".
Moreover, interpreting the results of the OSC analyses also requires knowledge of the properties of the methods they (and others) use.
For instance, how sensitive are these analyses? 
The irony is that answering this second question, in some ways, requires an answer to the former: how do we define replication?

In this paper, we clarify subjective notions about replication using a meta-analytic framework. 
Given this definition, we then evaluate various analysis methods to determine their properties.

# Defining Successful Replication: Meta-analytic Perspectives

Meta-analysis represents the "results" of an experiment by an effect parameter, denoted $\theta$.
This "true" effect reflects what the researcher would observe if they had an infinite sample size.
Effect parameters are typically on a standard scale, such as a standardized mean difference, Fisher-transformed correlation coefficient, or log-odds ratio.

In the context of replication studies, we have at least $k \geq 2$ such findings, so let $\theta_i$ describe the results of the $i$th study.
For the RPP and RPE programs, this invovles $\theta_1$ for the original study and $\theta_2$ for the replicate.
However, we do not actually observe $\theta_1$ and $\theta_2$, and instead we estimate them by $T_1$ and $T_2$, and these estimates have variances $v_1$ and $v_2$.
The estimates and their variances are what get reported in typical statistical analyses of experiments.

We argue that one useful definition of replication involves the similarity not of estimates $T_i$, but of actual true effects $\theta_i$.
The $T_i$ might differ because of both differences in the underlying true effects, but also due to random chance.
The $\theta_i$ might differ, however, because studies sample participants from different populations, or if there are differences in experimental contexts and conditions between experiments.
Thus, it may make sense to define replication in terms of differences in experiments rather than statistical noise.
Note that there is a connection to the potential outcomes framework in causal inference; for discussion see Wong and Steiner (2017).

In that vein, $\theta_1 = \theta_2$ would be a convenient starting point for defining replication.
However, there is evidence that even in the face of strong theory and sound scientific practice, that studies may still get slightly different results.
Thus, a more practical definition would say studies replicate if $\theta_1 \approx \theta_2$.
For further discussion on this distinction see Hedges and Schauer (2018).

The key distinction here, is that the definition of replication involves the parameters $\theta_i$. 
Analyses of replication involve the estimates $T_i$ and their variances $v_i$, which we use make inferences about how similar the $\theta_i$ are.
This paradigm allows us to evaluate properties of analysis methods.

For this paper, we interpret the OSC analyses in the tradition of statistical hypothesis testing.
Each analysis can be seen of a test of the null hypothesis $H_0: \theta_1 = \theta_2$, the metric used can be seen as a test statistic, and determinations that a finding failed to replicate correspond to a rejection of this null hypothesis.
Note that this is not the only structure we might infer here, and is not even the only way to operationalize these methods as hypothesis tests.