### 2018-05-30 effect sizes/effect_sizes.Rmd
Started effect size analyses

---
title: "Effect Sizes"
output: pdf_document
---

# Confidence Interval Comparison

Another approach used by the OSC and others to determine if a replication was successful involves confidence intervals. 
They determined that studies replicated if the estimate from the initial experiment $T_1$ is contained in a 95\% confidence interval for the replication effect parameter $\theta_2$. 
The logic behind this can be seen in two different ways. 
On the one hand, it can be conceived of as determining if the effect parameter in the replicate $\theta_2$ is different from the estimate of the initial study $T_1$. 
Note that they will be equal with probability zero under the model.

Another way to think about this analysis is that it tests whether the underlying effect parameters are the same.
To see this, note that $T_1$ being contained in the 95\% confidence interval for $\theta_2$ is equivalent to 
\[
S = |T_1 - T_2|/\sqrt{v_2} \leq 1.96
\]
This is conceptually identical to testing a null hypothesis of $H_0: \theta_1 = \theta_2$ with test statistic $S$ being compared to a standard normal distribution.
However, under $H_0$, $S$ does not follow a standard normal distribution.