### 2018-05-30 Ttest-pvals/writeup.Rmd
Write up for t-test of pvalues start

---
title: "Write Up: OSF T-Test"
author: "Mena Whalen"
output: pdf_document
---

1. OSF Article
A) steps list of what they did
B) why they did it

2. Theory
why is it bad
case 1 to case 2
3. Results sims
explain table and results

### OSF Article
 In the OSF article the authors used a paired t-test to compare the p-values from the original study with its replicate study for the 100 studies. The authors omit one case where the p-values were unavailable for both original and replicate. The test statistic for the paired t-test looks like $t=\frac{\bar{X}_{Diff}}{\frac{SD_{Diff}}{\sqrt k}}$ where $\bar{X}_{Diff}$ is the mean of all the differences between original and replicate p-values, $SD_{Diff}$ is the standard deviation of those differences from the mean, and $k$ is the number of pairs. The t-test is testing the hypothesis that these two groups are different from one another, meaning that the p-value from the first experiment done is the same p-value as the replicate study done. A paired t-test is used in this case since each study is being compared with its replicate so they are assumed to be controlling the particular experiment being conducted within each of the different groups, original and replicate.

 For the 99 pairs available the dependent t-test was done, finding a mean difference to be -0.2738438 showing that the replicate p-values were larger than the original. This leads to a test statistic of -8.2068 and less than 0.0001 p-value. This would lead to saying that the pairs are not the same or that their difference is equal to 0, thus the original studies p-values are not the same and the replicate studies p-values. Leading back to that the original studys' estimate is not the same as the replicates' estimate.

### Theory
#pvalues what goes into it
#power
#hypothesis meaning (mult. hypothesis)
#case 1 and case 2 what happens
 Using the set up provided with $T_i \sim N(\theta_i, v_i)$, if a study is significant, meaning p-value is less than or equal to 0.05, then $\frac{|T_i|}{\sqrt{v_i}} \geq 1.96$ if testing $H_0:\theta_i =0$. We assume that the variance is known for each study and is $\frac{4}{n}$ where $n$ is the sample size. This means that the test statistic used for determining the p-value is dependent on the size of the estimate and the sample size. This p-value is a random variable coming from the given data, the distribution of which is dependent on the power of the test. The power to detect if $\theta_i$ is statistically different from 0 depends on the estimate $T_i$ and the sample size $n$. Given that each original study and its replicate study has possibly different estimates and sample sizes to calculate its p-value means that the groups could have different statistical power for each test.

 When comparing the p-values of two paired groups using the t-test incorporates each cases' estimate and the sample size of each test. In a single example, say study 1, there would be $p_1$ the p-value from the orginal and $p_2$ the p-value from the replicate study. This is equivalent to $p_1=\Phi^{-1}\Big(\frac{|T_1|}{\sqrt{v_1}}\Big)$ and similar can be shown for $p_2$, so subtracting the two leads to $\Phi^{-1}\Big(\frac{|T_1|}{\sqrt{v_1}}\Big)-\Phi^{-1}\Big(\frac{|T_2|}{\sqrt{v_2}}\Big)$. This is dependent on the power of each statistical test using the $T$'s and $n$'s for each group. This scaled for k studies, the mean of the paired difference of p-values would be 
\begin{align}
\frac{1}{k} \sum_i ^k \Big[ \Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big)-\Phi^{-1}\Big(\frac{|T_i|}{\sqrt{v_i}}\Big) \Big]
\end{align}
This mean takes into account both the estimate and the sample size, thus the power to detect if the p-values are different from one another depends upon the power of both the original and the replicate studies.

 In the current situation the t-test is testing if the group of original studies are similar to the group of replicate studies over all studies using p-values to assess similarity. When examining all studies together this takes away from a single study being replicated. Each study is dependent upon the power of both original and replicate being able to detect an effect. This is not the same thing as testing if an individual study did replicate which would relate back in hypothesis test form to 
 \begin{align}
 H_0: \theta_1 = \theta_2 \hspace{1cm} H_A: \theta_1 \neq \theta_2
 \end{align}
where $\theta_1$ is the parameter from the original and $\theta_2$ from the replicate.