### 2020-05-19 Fishers_method/Fishers method tables and plots.Rmd

---
title: "Fisher's method tables and plots"
author: "Katie Fitzgerald"
date: "1/15/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Table: RPE- Power of Fisher's method given large sample sizes for varying $\delta$ and true # of non-null effects (i.e. "Best-case scenario")
```{r, echo=FALSE, warning= FALSE, message = FALSE}
library(knitr)
library(tidyverse)
RPP_RPE_power_results <- readRDS("./RPP_RPE_power_results.RDS")
table <- spread(RPP_RPE_power_results, key = delta, value = power)
table_rpp <- table[table$study == "RPP", -c(1,3, 4)]
table_rpe <- table[table$study == "RPE", -c(1,3, 4)]
kable(table_rpe, format="markdown", digits=4, align=c('l','r','r','r'),row.names = FALSE, col.names=c("# of false negatives","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="RPE: Power to reject H0 with large n")
```

Table: RPP- Power of Fisher's method given large sample sizes for varying $\delta$ and true # of non-null effects (i.e. "Best-case scenario")
```{r, echo=FALSE, warning= FALSE, message = FALSE}
kable(table_rpp, format="markdown", digits=4, align=c('l','r','r','r'),row.names = FALSE, col.names=c("# of false negatives","$\\delta=0.2$","$\\delta=0.5$","$\\delta=0.8$"),caption="RPP: Power to reject H0 with large n")
```

```{r echo = FALSE, warning = FALSE, message = FALSE , fig.cap = "Power of Fisher's method, best-case scenario using RPE and RPP sample sizes \\label{fig:sig:plot_all}"}
library(ggplot2)
library(wesanderson) # color palettes
zs1 = wes_palette("Zissou1")
dj1 = wes_palette("Darjeeling1")
dj2 = wes_palette("Darjeeling2")

# standard theme for plots
stdtheme = theme(panel.grid.minor = element_blank(),
                 axis.ticks = element_blank(),
                 axis.text = element_text(size = 16),
                 axis.title = element_text(size = 18),
                 legend.title = element_text(size = 16),
                 legend.text = element_text(size = 16),
                 legend.text.align = 0)

study_labels <- c( 'RPE' = "RPE, r = 7", 'RPP' = "RPP, r = 64")

plot_all <- ggplot(RPP_RPE_power_results, 
                   aes(x = prop, y = power, linetype = factor(delta)))+
  facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
  stdtheme +
  theme(strip.text = element_text(size = 16))+
  scale_x_continuous(name = "Proportion of false negatives") +
  scale_y_continuous(name = "Power of Fisher's method")+
  labs(linetype = expression(delta))+
  geom_line()
plot_all

plot_worst <- ggplot(RPP_RPE_worst, aes(x = prop, y = power, linetype = factor(delta)))+
  facet_grid(rows = vars(study), labeller = labeller(study = study_labels)) +
  stdtheme +
  theme(strip.text = element_text(size = 16))+
  scale_x_continuous(name = "Proportion of false negatives") +
  scale_y_continuous(name = "Power of Fisher's method")+
  labs(linetype = expression(delta))+
  geom_line()
plot_worst
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "Power of Fisher's method, general case \\label{fig:sig:plot_general_case}"}

###CREATE PLOT FOR GENERAL SETTING####
k10 <- readRDS("./k10_data.RDS")
k50 <- readRDS("./k50_data.RDS")
k100 <- readRDS("./k100_data.RDS")

n_labels <- c('25' = "n=25", '50' = "n=50", '75' = "n=75", '100'= "n=100")
k_labels <- c( '10' = "k=10", '50' = "k=50", '100' = "k=100")

k10$prop <- k10$M/k10$k
k50$prop <- k50$M/k50$k
k100$prop <- k100$M/k100$k

all <- rbind(k10,k50,k100)
plot_general_case <- ggplot(all, aes(x = prop, y = power, linetype = factor(delta))) +
  #facet_grid(k~n, labeller = labeller(n = n_labels, k = k_labels)) +
  stdtheme + 
  theme(axis.text = element_text(size = 6), strip.text = element_text(size = 16)) +
  scale_x_continuous(name = "Proportion of false negatives" ) +
  scale_y_continuous(name = "Power") +
  labs(linetype = expression(delta))+
  geom_line()

plot_general_case

power_80_k10 <- dplyr::filter(all, k == 10 & ((delta == .8 & n == 25 ) | (delta == 0.5 & (n == 50 | n == 75)) | (delta == 0.2 & n == 100)))

```